---
title: "Introduction to R Markdown"
author: "Christopher Swader"
date: "`r Sys.Date()`"
output:
  powerpoint_presentation: default
  #word_document  
---

# Introduction

Welcome to the R Markdown workshop! Today, we'll learn how to create documents and presentations using R Markdown.

# What is R Markdown?

- **R Markdown** is a file format for making dynamic documents with R.
- It combines plain text with embedded R code.
- You can create reports, presentations, dashboards, and more. It allows the results of your R code, analyses to directly enter the final document. 

# Installing R Markdown

To install R Markdown, use the following command in R:
```{r}
library("rmarkdown")
#also load any other libraries that are used in this file!!
library(ggplot2)
```
Of course install the package first if needed. 

# Creating a New R Markdown File
In RStudio:

Go to File > New File > R Markdown...
Select Document or Presentation
Choose your desired output format (e.g., HTML, PDF, Word, ioslides, Slidy, Beamer, PowerPoint)

# Basic Structure of an R Markdown File
This is the YAML header, which goes at the top of your rmd file. 
```{r, eval=FALSE}


---
title: "My Document"
author: "Your Name"
date: "2024-09-21"
output: html_document
---
```
# Code chunks
- Code chunks are sections of your document that contain R code.
- They start with three backticks and `{r}` and end with three backticks.
-This is an example of an R Markdown code chunk. *The results* of the chunk will automatically enter the final document, unles you set {r, eval=FALSE}.

```{r}
# You can click the green arrow to the right of the chunk in Rstudio to see the result.
summary(cars)
```


# Your turn
Task: Write a function to calculate the mean of a numeric vector cars$speed. Add it to the next slide within a code chunk.

# Your solution here
```{r}
# Your vector

```

# Adding Plots
You can include plots in your R Markdown document using code chunks.

Here is a simple (non-ggplot) Example:
```{r}
# This is a code chunk

plot(mtcars$mpg, mtcars$hp)
```


# Your turn
Task: Create a scatter plot of the mtcars dataset showing miles-per-gallon *mpg* plotted against horsepower *hp* and enter it in the next slide, code chunk. Use ggplot!

# Your solution here
Example:
```{r}
# This is a code chunk

```

# A solution
```{r}
# This is a code chunk
ggplot(data = mtcars, aes(x=hp, y=mpg))+geom_point()


```

# Adapting the text

## Header level
## There is always a space between the pound sign and the text afterwards. A single pound sign means new slide.
*italics*, 
**bold**

# Customizing Your Document
You can customize the appearance of your document using YAMLheader metadata, depending on the output type.
Add themes, highlight styles, and more.
You can also change the output type. 
-Let's change this to a word document Change this at the top of this rmarkdown file and knit.

```{r, eval=FALSE}


---
title: "My Document"
output:
word_document 
---
```


# Your turn
* Create a new slide with text introducing the first five lines (use head() function) of the flfp_agg dataset. 
* Change the output to power point slides again in the header
* Knit the new rmd, adjust any slides that are needed, and save your powerpoint slides.


